Response to Referees

### 2025-06-05 ms/ms.Rmd
address RC1.5: radiometric cps/concentration

>>>>

Although radiometric data must be in units of counts per second to model attenuation directly [@olearyDigitalSoilMapping2022], we used the radiometric data as provided to us: in units of concentration for potassium, thorium, and uranium (converted from counts per second by scalar calibration factors).
The monotonic transformation between counts per second and concentration does not affect the tree-based machine learning algorithm that we used to model peat depth [@hastieElementsStatisticalLearning2009].

### 2025-06-05 ms/ms.tex
address RC1.5: radiometric cps/concentration

>>>>

Although radiometric data must be in units of counts per second to model attenuation directly \citep{olearyDigitalSoilMapping2022}, we used the radiometric data as provided to us: in units of concentration for potassium, thorium, and uranium (converted from counts per second by scalar calibration factors).
The monotonic transformation between counts per second and concentration does not affect the tree-based machine learning algorithm that we used to model peat depth \citep{hastieElementsStatisticalLearning2009}.

### 2025-06-05 ms/submission/AuthorComments-egusphere-2025-1046.qmd
address R2 Technical corrections - depths table
address RC1.5: radiometric cps/concentration

---
title: "Response to Referees"
author: "Julien Vollering, on behalf of all co-authors"
format:
  html:
    embed-resources: true
toc: true
---

@@

@@

>>>>

---
title: "Response to Referees"
author: "Julien Vollering, on behalf of all co-authors"
format:
  html:
    embed-resources: true
toc: true
bibliography: ../ms.bib
---

@@

Thank you to the referee for pointing out this concern and providing a thorough explanation of their reasoning. We appreciate the references, which we have consulted.

After careful consideration, we disagree that the radiometrics variables we used are not valid for predicting peat depth. In particular, we assert that converting the concentrations we use to counts per second -- as the referee suggests -- would produce identical results. That is because the method we use to predict peat depth (Random Forest) is insensitive to any monotonic transformation of the predictors, like the scalar conversion between concentration and counts per second [see Table 10.1 on p.351 in @hastieElementsStatisticalLearning2009]. The Random Forest algorithm is based on decision trees, which partition the predictor space into regions based on the values of the predictors. As a result, any monotonic transformation of the predictors will not change the partitioning of the predictor space, and therefore will not change the predictions made by the model [see Fig. 9.2 on p.306 and Algorithm 15.1 on p.588 in @hastieElementsStatisticalLearning2009]. Appendix B3 in @baranwalHelicopterborneMagneticElectromagnetic2013 and Appendix A2 in @ofstadHelicopterborneMagneticRadiometric2015 confirm that the concentration values we use in our models are simply the counts per second multiplied by a constant conversion factor.

To try to avoid confusion on this point, we have added a short explanation in the Methods section that the distinction between counts per second and concentration is not relevant for our analysis and it does not affect the predictive power of the radiometric variables in our models.

[Although radiometric data must be in units of counts per second to model attenuation directly [@olearyDigitalSoilMapping2022], we used the radiometric data as provided to us: in units of concentration for potassium, thorium, and uranium (converted from counts per second by scalar calibration factors).
The monotonic transformation between counts per second and concentration does not affect the tree-based machine learning algorithm that we used to model peat depth [@hastieElementsStatisticalLearning2009].]{.added} 
:::

@@

We have moved the table after its first mention and adjusted the table caption to clarify that the percentages represent the proportion of the total number of 10 m cells across the stratifications. We have also added standard deviations to the mean depths, in accordance with comment 2.5.
